Pulsar Classification using Deep Neural Networks

### 2018-04-23 paper/paper.tex
Mostly done with the paper. Finalizing, spellchecking, adding any other citations.

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}

@@

\subsection{Data Format}
  After determining the dataset and the framework to be utilized it is 
  necessary to figure out how to read data from the dataset. A modular approach 
  was taken with this project. Users will be able to select a comma-separated 
  values (CSV) file to load in. By specifying a class to contain the CSV data it 
  is possible to run multiple datasets while making only minor modifications to 
  the code. The format is also much easier to read as a result of the class 
  format.

@@

  For that endevour scikit-learn was used, specifically the 
  $train\_test\_split()$ method. As is standard for classification problems the 
  data was segmented into 80\% training and 20\% testing data. It was then 
  necessary to place the $class$ label into a data structure for later. For both 
  the train and test data, associated labels are passed, along with the data, 
  as tuples. The psuedocode for this section can be found under Algorithm 
  \ref{alg:input_fn}.

  \begin{algorithm}
    \caption{$input\_fn()$}
    \label{alg:input_fn}
    \begin{algorithmic}
      \STATE $data \gets CSV$\\
      \STATE $test, train \gets train\_test\_split(data)$
      \STATE $test\_label \gets test.pop(class)$
      \STATE $train\_label \gets train.pop(class)$
      \RETURN $(train, train\_label), (test,test\_label)$
    \end{algorithmic}
  \end{algorithm}

\section{Retrieval of style files}

The style files for NIPS and other conference information are
available on the World Wide Web at
\begin{center}
  \url{http://www.nips.cc/}
\end{center}
The file \verb+nips_2017.pdf+ contains these instructions and
illustrates the various formatting requirements your NIPS paper must
satisfy.

The only supported style file for NIPS 2017 is \verb+nips_2017.sty+,
rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{}
  2.09, Microsoft Word, and RTF are no longer supported!}

The new \LaTeX{} style file contains two optional arguments:
\verb+final+, which creates a camera-ready copy, and \verb+nonatbib+,
which will not load the \verb+natbib+ package for you in case of
package clash.

At submission time, please omit the \verb+final+ option. This will
anonymize your submission and add line numbers to aid review.  Please
do \emph{not} refer to these line numbers in your paper as they will
be removed during generation of camera-ready copies.

The file \verb+nips_2017.tex+ may be used as a ``shell'' for writing
your paper. All you have to do is replace the author, title, abstract,
and text of the paper with your own.

The formatting instructions contained in these style files are
summarized in Sections \ref{gen_inst}, \ref{headings}, and
\ref{others} below.

@@

\end{document}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{pdfpages}       % Because LaTeX doesn't take SVGs...
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{natbib}

@@

\subsection{Main}
  Main.py is where the project is ran from. There are several hyperparameters 
  that a user can pick, beginning with $\_TEST\_PERCENTAGE$, which sets the 
  percentage of comma-separated values (CSV) data to be utilized for testing. 
  Industry standard is 20\%, so that is the default setting. The training data 
  $\_TRAIN$ is, as a result, 80\% of the CSV data. 
  % 
  \[\mathcal{\_TRAIN}=\_NUM\_ELEMENTS * ( 1 - \_TEST\_PERCENTAGE )\]
  % 
  It was necessary to hard-code in the number of elements in the CSV file because there was trouble when trying to run update on the CSV class due to 
  an incomplete understanding of python abilities such as $class.update()$.

  $\_BATCH\_SIZE$ is set as 
  %
  \[\mathcal{\_BATCH\_SIZE}=max\{\_TRAIN*0.1, 1\}\]
  %
  This is part of ensuring the training does not get stuck in local minima. \_MODEL is then set as whichever optimizer the user would like, in this case, 
  RMSPropOptimizer was selected. A CSV class is created based on the dataset to 
  be classified.

  Hyperparameter choices are part of the art of machine learning but in this 
  project it was not the focus to make perfect hyperparameters. That being 
  said, that was explored briefly for this project as seen in section 
  \ref{sec:hyperparams}. $\_NODES\_PER\_LAYER$ was one of such hyperparameters, 
  as was $\_LAYERS$, and $\_HIDDEN\_LAYERS$. The final implementation of these 
  parameters are as follows:
  %
  \[\mathcal{\_NODES\_PER\_LAYER}=(num\_features - 1) * 1.5\]
  \[\mathcal{\_LAYERS}=max\{sqrt(\_NODES\_PER\_LAYER), 2\}\]
  \[\mathcal{\_HIDDEN\_LAYERS}=[\_NODES\_PER\_LAYER] * \_LAYERS\]
  %
  This lead to 3 hidden layers with 12 nodes each for the \href{https://archive.ics.uci.edu/ml/datasets/HTRU2}{HTRU\_2 Dataset}.

  After all parameters have been setup main gets run. Users can select the 
  number of runs to perform as an argument when they begin the program e.g. 
  %
  \[\textnormal{python\ main.py \textit{X}}\]  
  %
  The program will then run the classifier \textit{X} times. After finishing 
  the program will output classification accuracy of the given dataset.

\subsection{Data format}
  After determining the dataset and the framework to be utilized it is 
  necessary to figure out how to read data from the dataset. A modular approach 
  was taken with this project. Users will be able to select a CSV file to load 
  in. By specifying a class to contain the CSV data it is possible to run 
  multiple datasets while making only minor modifications to the code. The 
  format is also much easier to read as a result of the class format.

@@

  For that endevour scikit-learn was used, specifically the 
  $train\_test\_split()$ method. As is standard for classification problems the 
  data was segmented into 80\% training and 20\% testing data. It was then 
  necessary to place the $class$ label into a data structure for later. For 
  both the train and test data, associated labels are passed, along with the 
  data, as tuples. The psuedocode for this section can be found under Algorithm 
  \ref{alg:input_fn}.
  %
  \begin{algorithm}
    \caption{$input\_fn()$}
    \label{alg:input_fn}
    \begin{algorithmic}
      \STATE $data \gets CSV$\\
      \STATE $test, train \gets train\_test\_split(data)$
      \STATE $test\_label \gets test.pop(class)$
      \STATE $train\_label \gets train.pop(class)$
      \RETURN $(train, train\_label), (test,test\_label)$
    \end{algorithmic}
  \end{algorithm}
  %
\subsection{TensorFlow estimator}
  TensorFlow has built in estimators such as:
  %
  \begin{itemize}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/DNNClassifier}{DNNClassifier}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/BaselineClassifier}{BaselineClassifier}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/BaselineRegressor}{BaselineRegressor}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/DNNLinearCombinedClassifier}{DNNLinearCombinedClassifier}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/LinearClassifier}{LinearClassifier}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/LinearRegressor}{LinearRegressor}
  \end{itemize}
  %
  You also have the ability to build your own estimator, or classifier, and 
  this is the route that was selected for this project. The custom estimator 
  selected utilizes the $model$ specified at the start of the program, a 
  $model\_dir$ for TensorBoard output which will be discussed in section 
  \ref{sec:tensorboard}, and the $params$ which consist of $feature\_columns$, 
  $hidden\_units$, and $n\_classes$.

  Estimator also requires \href{https://www.tensorflow.org/versions/r1.2/api_docs/python/tf/estimator/ModeKeys}{three modes} depending on the goal:
  %
  \begin{itemize}
    \item ModeKeys.TRAIN
    \item ModeKeys.EVAL
    \item ModeKeys.PREDICT
  \end{itemize}
  %
  Lastly, the estimator requires a return of $EstimatorSpec$. Documentation for 
  the EstimatorSpec can be found on \href{https://www.tensorflow.org/api_docs/python/tf/estimator/EstimatorSpec}{TensorFlow's website}. It is important to note that each of the different $ModeKeys$ requires a different set of parameters for the returned $EstimatorSpec$.

\subsection{Training}
  TensorFlow offers the ability to train estimators by calling a $train()$ 
  method. For this paper the $input\_fn$ and $steps$ arguments are set by a 
  function call to $train\_input\_fn$ and $csv.num\_examples['train']$ 
  respectively. This allows the data to be segmented for training.

  The model that was instantiated during estimator creation is detailed under
  $\textnormal{model.model\_RMSProp}$. For each of the layers that was passed 
  to it under $params['hidden\_units']$ there is a fully connected layer built 
  with selu activation. \citet{selu} found with the proper math you can obtain 
  self-normalizing exponential linear units, meaning they don't require batch 
  normalization and that is why they were chosen over all of the other 
  activation functions. There is also the final fully connected layer which has 
  no activation layer. This final layer aids in determining the classification.

  During training ModeKeys.TRAIN is called, which specifies RMSProp gradient 
  descent as our optimizer and then calls $minimize()$. The $minimize()$ method 
  will automatically compute the gradients and apply them to the variables, but 
  if you need to tune the network there are additional steps that can be taken 
  instead:
  %
  \begin{enumerate}
    \item Compute the gradients with $compute\_gradients()$
    \item Process the gradients as you wish
    \item Apply the processed gradients with $apply\_gradients()$
  \end{enumerate}
  %
  Further documentation on this process can be found on TensorFlow's website 
  under the \href{https://www.tensorflow.org/api_docs/python/tf/train/Optimizer}
  {class Optimizer}. It was not necessary to have that level of fidelity with 
  this particular dataset, so the base $minimize()$ was all that was necessary.

  With $minimize()$ minimizing the loss we are able to make progress towards a 
  classification. This dataset in particular minimizes loss to near-zero after 
  just the first episode of 100 training steps as can be seen in Figure \ref{img:loss}.

\subsection{Hyperparameters}\label{sec:hyperparams}
  Hyperparameters are normally very important but in this dataset it was 
  discovered that they had very little impact on the efficacy of 
  this classifier.

  An example of hyperparameter randomization can be found under:
  %
  \[\textnormal{python\ hyperparams.py}\]
  %
  That file will automatically generate randomized hyperparameters. During 
  testing of this project various hyperparameters were tested with no 
  significant improvements over the initial parameters set so the associated 
  code was removed from primary operation of the classifier. The file described 
  would allow for random testing of hyperparameters across any number of 
  computers. \citet{hypparam} found random search of hyperparameters to be superior to grid search and was the method used while exploring 
  hyperparameter tuning during this project.

\subsection{TensorBoard}\label{sec:tensorboard}
  One of the other reasons for utilizing TensorFlow is the TensorBoard project. 
  TensorBoard allows for visualization of various parameters as defined by the 
  user. Figure \ref{img:acc_smooth} and \ref{img:loss_smooth} are the same results as shown in Figure \ref{img:accuracy} and \ref{img:loss} 
  respectively. The smooothing helps in understanding what may be going on, 
  especially in more turbulant graphs. Another benefit is being able to 
  visualize the graph structure. To use TensorBoard on this project ensure it 
  was installed with TensorFlow and type 
  \[\textnormal{tensorboard\ -\--logdir\ model}\]
  and go to the link it provides (by ctrl clicking or copy/paste).

  \begin{figure}[H]
  \caption{Smoothed Accuracy}\label{img:acc_smooth}
  \begin{center}
  \includegraphics[scale=0.65]{../images/acc}
  \end{center}
  \end{figure}
  %
  \begin{figure}[H]
  \caption{Smoothed Loss}\label{img:loss_smooth}
  \begin{center}
  \includegraphics[scale=0.65]{../images/loss}
  \end{center}
  \end{figure}

\newpage
\section{Results}\label{sec:results}
  RMSProp gradient descent with 0.001 learning rate, performing batch runs on 
  10\% of the training samples every 100 steps produced an accuracy of 97.8\% 
  averaged over 20 tests. These were taken with 80\% train and 20\% test 
  samples. Other hyperparameters were three fully connected layers with 12 
  nodes each using selu activation.

  \begin{figure}[H]
  \caption{Accuracy}\label{img:term_acc}
  \begin{center}
  \includegraphics[scale=0.75]{../images/accuracy_percent_small}
  \end{center}
  \end{figure}
  %
  \begin{figure}[H]
  \caption{Accuracy}\label{img:accuracy}
  \begin{center}
  \includegraphics[scale=0.75]{../images/Accuracy}
  \end{center}
  \end{figure}
  %
  \begin{figure}[H]
  \caption{Loss}\label{img:loss}
  \begin{center}
  \includegraphics[scale=0.75]{../images/Loss}
  \end{center}
  \end{figure}

\section{Conclusion}

@@

% [4] Bergstra, J.\ \& Bengio, J.\ (2012) Random Search for Hyper-Parameter 
% Optimization. {\it Journal of Machine Learning Research 13} pp.\ 281-305.

\bibliographystyle{plainnat}
\bibliography{bib}
\end{document}
